### 2020-09-02 manual_distribution_max.Rmd
mean null updated with scripts all in mean_null.R

---
title: "Untitled"
author: "Sayani Gupta"
date: "28/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why normalise?

Suppose that $X_1$, $X_2$ , ... , $X_n$ are i.i.d. random variables with expected values $E(X_i) = \mu < \infty$ and variance $Var(X_i) = \sigma^2 < \infty$. Let 
$Y = max(X_1, X_2, \dots, X_n)$.

Let $F_X(x)$ be the common distribution of the variables $X_i$ and let $F_Y(y)$ be the
corresponding distribution of $Y$. $F_Y(y)$ could be obtained from $F_X(x)$ simply by using:
$F_Y(y) = P[(X_1 ≤ y)∩(X_2 ≤ y) ∩...∩ (X_n ≤ y)] = {F_X(y)}^n$. For large $n$, the distribution of $Y$ approaches a standard shape, which does not depend on $F_X$.But what about the case when $n$ is not large enough. Then the distribution of maximum will indeed depend on maximum is taken for random variables from which distributions. If $F_X(x)$ is the CDF of $X$, then $F_Y(y) = {F_x(y)}^n$. If $\Phi$ nd $\phi$ are the cdf and pdf of a standard normal distribution, then
$f_Y(y) = n{\Phi(y)}^{n-1}\phi(y)$, which depends on $n$. Hence, we are trying to normalise for $n$, since the distribution of maximum depends on $n$.As $n$ grows, we  can see the right tail growing, which implies that the probability that we will get a higher maximum is more. Now, for large $n$, we used EVT to normalise for $n$, that is, we brought them to the same scale without distorting the range of the distribution. But in our case, we will mostly have small $n$. It is important to ensure that they have the same mean and variation, for being able to compare the maximum value across $n$.

## Mean and standard deviation of the distribution of maximum for small n

Analytically,
sample mean: $ \frac{\sum(Y_i)}{n}$
throeretical mean: $ \frac{\int_i(y_if_Y(y_i))}{n}$