### 2020-12-07 simulations/objectives.Rmd
renv update

---
title: "Simulation objectives"
author: "Sayani Gupta"
date: "01/12/2020"
output: pdf_document
---

This file contains objectives of the simulations and description of parameters and files or folders in simulations/Results/ with corresponding objective in parenthesis.

## Objectives:

 (1A) to understand how the measure (raw or normalised) changes across different levels of facets and x-axis for fixed $nsim$, $nperm$ and $\lambda$ in null case. Original values are **not scaled**, sim_table has just nx and nfacet. Description of parameters: 
 - $nfacet = nx = c(2, 3, 5, 7, 14, 20, 31, 50)$
 - $nsim = 200$
 - $nperm = 100$
 - $lambda = 0.67$
 - $sim_dist = normal(5, 10)$
 - $weights (dist_{within-facet}, dist_{across-facet}) = (\lambda, 1/\lambda)$

(1B)  Original values are **scaled** and other parameters look exactly like (A)

(2)  Original values are **scaled** and $weights (dist_{within-facet}, dist_{across-facet}) = (\lambda, 1 - \lambda)$, other parameters like (A), for $lambda = seq(0.1, 0.9, 0.05)$

## Description of files

Folder raw/

- Files ending with _dist.rds : ran from null/job-raw.sh (A)
- Files ending with _pairwise_max_dist.rds: ran from null/job-pairwise_max.sh (A)
- Folder data:  ran from null/job-pairwise_data.sh (A)

Folder norm/

- Files ending with _dist.rds: run from null/job-norm.sh (A)