### 2021-01-27 scalar_transformation.Rmd
scalar transformation document update

---
title: "Choosing an appropriate scalar transformation to normalise wpd"
author: "Sayani Gupta"
date: "28/01/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE)
knitr::read_chunk("example_transformations.R")
library(tidyverse)
library(readr)
library(ggplot2)
```

 - Data presented

Observations are generated from a N(0,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high. Each combination is being referred to as a _panel_. That is, data is being generated for each of the panels $\{nx = 2, nfacet = 2\}, \{nx = 2, nfacet = 3\}, \{nx = 2, nfacet = 5\},  \dots, \{nx = 50, nfacet = 31\}, \{nx = 50, nfacet = 50\}$. For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories. That is, if we consider the panel  $\{nx = 2, nfacet = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$. The values of $\lambda$ is set to $0.67$ and values of raw wpd is obtained.

 - How the distribution of the raw wpd looks across nfacets and nx?

 Both shape and scale of the distribution changes for different nx and nfacet categories.

```{r read}
N01 <- read_rds("simulations/raw/null_design_quantrans/data-agg/all_data_wpd_N01.rds")
```

```{r raw}
N01 %>% 
  ggplot(aes(x = value)) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  scale_x_continuous(breaks = scales::breaks_extended(3)) + 
  xlab("raw wpd")
```

 - Plot the values of wpd against nx*nfacet to see the rough relationship

```{r}
N01 %>% 
  ggplot(aes(x=as.factor(nx*nfacet), y = value)) +
  geom_boxplot() + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd")

```
 - Plot the values of wpd against nx*nfacet to see if the same relationship holds for different nx and nfacet relationships

Looks like for all the variations, there is a quadratic relationship between wpd values and nx*nfacet

```{r}
every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}
```

```{r}
N01 %>% 
  mutate(comp_lev = if_else(nx<nfacet, "nx<nfacet",
                            if_else(nx>nfacet,
                               "nx>nfacet", "nx=nfacet"))) %>% 
  ggplot(aes(x=as.factor(nx*nfacet), y = value)) +
  geom_boxplot() + facet_wrap(~comp_lev) + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + xlab("nx*nfacet") + ylab("raw wpd") + scale_x_discrete(breaks = every_nth(n = 5))
```

 - Attempt to linearize it with sqrt, sqrt(sqrt) and log (sqrt(sqrt))

 The transformation log (sqrt(sqrt())) on nx*nfacet finally makes it approximately linear.

```{r}
N01 %>% 
  mutate(comp_lev = if_else(nx<nfacet, "nx<nfacet",
                            if_else(nx>nfacet,
                                    "nx>nfacet", "nx=nfacet"))) %>% 
  ggplot(aes(x=sqrt(nx*nfacet), y = value)) +
  geom_point() + facet_wrap(~comp_lev) + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + ylab("raw wpd")

N01 %>% 
  mutate(comp_lev = if_else(nx<nfacet, "nx<nfacet",
                            if_else(nx>nfacet,
                                    "nx>nfacet", "nx=nfacet"))) %>% 
  ggplot(aes(x=sqrt(sqrt(nx*nfacet)), y = value)) +
  geom_point() + facet_wrap(~comp_lev) + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + ylab("raw wpd")

N01 %>% 
  mutate(comp_lev = if_else(nx<nfacet, "nx<nfacet",
                            if_else(nx>nfacet,
                                    "nx>nfacet", "nx=nfacet"))) %>% 
  ggplot(aes(x=log(sqrt(sqrt(nx*nfacet))), y = value)) +
  geom_point() + facet_wrap(~comp_lev) + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + ylab("raw wpd")

```

 - Distribution after linearizing it

The distribution of $wpd_{norm} = log(sqrt(sqrt(nx*nfacet)))/wpd$ is plotted. The shape and spread look similar but location is shifting to the right.

(If we define $wpd_{norm}$ as the inverse of it, the values become too small and the distribution too skewed. Hence the inverse of it is considered.)

<!-- even after doing Box-cox transformation to up-weight the lower values of nx*nfacet and down-weigh higher values -->

```{r}
 N01 %>% 
  ggplot(aes(x = forecast::BoxCox(log(sqrt(sqrt(nx*nfacet)))/value,
             lambda = "auto"))) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd norm")

```

What we want is a transformation which will be constant and not linear to obtain similar locations for all panels.