### 2021-02-08 comparison_perm_scalar.Rmd
comparison of scalar and permutation transformation

---
title: "Compare permutation and scalar transformation approaches on simulated and real data"
author: "Sayani Gupta"
date: "08/02/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE)
knitr::read_chunk("example_transformations.R")
library(tidyverse)
library(readr)
library(ggplot2)
```

**Simulated data generation:** Observations are generated from a Gamma(2,1) distribution for each combination of $nx$ and $nfacet$ from the following sets: $nx = nfacet = \{2, 3, 5, 7, 14, 20, 31, 50\}$ to cover a wide range of levels from very low to moderately high. Each combination is being referred to as a _panel_. That is, data is being generated for each of the panels $\{nx = 2, nfacet = 2\}, \{nx = 2, nfacet = 3\}, \{nx = 2, nfacet = 5\},  \dots, \{nx = 50, nfacet = 31\}, \{nx = 50, nfacet = 50\}$. For each of the $64$ panels, $ntimes = 500$ observations are drawn for each combination of the categories. That is, if we consider the panel  $\{nx = 2, nfacet = 2\}$, $500$ observations are generated for each of the combination of categories from the panel, namely, $\{(1, 1), (1, 2), (2, 1), (2, 2)\}$. The values of $\lambda$ is set to $0.67$ and values of raw wpd $wpd_{raw}$ is obtained.

## Scalar transformation approach to normalisation

A log-linear model is fitted to see how the values of $wpd_{raw}$ changes with the values of $nx$ and $nfacet$. The model is of the form $$y = a+b*log(x) +e$$, where $y = median(wpd_{raw})$ and $x = nx*nfacet$. $wpd_norm$ is a transformation on $wpd_{raw}$ which should be designed to remove the effect of $nx*nfacet$ on $wpd_{raw}$ and thus is defined as follows:
$wpd_{norm} = wpd_{raw} - b*log(nx*nfacet)$

## Permutation approach to normalisation

The simulated data for each of the panels is permuted/shuffled $nperm = 100$ times and for each of those permutations $wpd_{norm}$ is computed as follows: $wpd_{norm} =  (wpd_{raw} - mean(wpd_{raw}))/sd(wpd_{raw})$
. This is done so that the distribution of the normalised measure $wpd_{norm}$ has the same mean and standard deviation across different nx and nfacet.

## Comparing the two approaches for simulated data

In Figure \ref{fig:dist-new-transform}, we see that the distribution of $wpd_{norm2}$ is more similar across nx and nfacets than $wpd_{norm1}$.

In Figure \ref{fig:dist-new-approx}, we see that the distribution of $wpd_{norm2}$ is similar when normalised with true or approximate estimates of a and b.

```{r dist-new-approx, echo = TRUE, fig.cap = " The distribution of $wpd_{norm2}$ is plotted for approximate estimate of a and b. The distributions are similar when plotted against true estimate of and b."}

G21 <- read_rds("simulations/raw/null_design_quantrans/data-agg/all_data_wpd_Gamma21.rds")

G21_median <- G21 %>% 
  group_by(nx*nfacet) %>% 
  summarise(actual = median(value))

# fit model median to log(nx*nfacet)
fit_lm2 <- lm(actual ~ poly(log(`nx * nfacet`) ,1, raw=TRUE), data = G21_median)

summary(fit_lm2)

intercept <- fit_lm2$coefficients[1]
slope <- fit_lm2$coefficients[2]

G21_compare <- G21 %>% 
  mutate(with_true_est = (value - slope*log(nx*nfacet)),

                          with_approx_est = (value - 0.0027*log(nx*nfacet))) %>% 
  pivot_longer(cols = 5:6,
               names_to = "type_estimate",
               values_to = "value_estimate")

G21_compare %>% 
  ggplot() +
  geom_density(aes(x = value_estimate, fill = type_estimate), alpha = 0.5) +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  scale_x_continuous(breaks = scales::breaks_extended(3)) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#0072B2", "#D55E00")) +
  xlab("wpd_norm2")
```

## Comparing the two approaches from real data