### 2021-11-03 paper/preamble.tex
Added preamble.tex

\setlength {\marginparwidth }{2cm}
\usepackage{mathtools,amssymb,booktabs,amsthm,todonotes,colortbl}
\def\mod{~\text{mod}~}
\usepackage{mathptmx}
\usepackage{caption,setspace}
\usepackage[section]{placeins}
\usepackage{longtable}
\DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true,font={stretch=1.2}}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true,font={stretch=1.2}}
\def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
\def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
\newcommand{\DC}[1]{#1}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}
\def\wpd{\text{\textit{wpd}}}
\def\wpdsub#1{\text{\textit{wpd}}_{\text{#1}}}
\def\Dnull{D_{\emptyset}}
\def\Df{D_{\!f}}
\def\Dfx{D_{\!fx}}
\def\Dx{D_{x}}
\def\nsub#1{n_{\text{#1}}}
\def\nx{n_x}
\def\nf{n_{\!f}}

%% LINE AND PAGE BREAKING
\sloppy
\clubpenalty = 10000
\widowpenalty = 10000
\brokenpenalty = 10000
\RequirePackage{microtype}